---
title: "Developing Data Products"
subtitle: "Week 3 Assignment"
author: "Yige"
date: "*Feb. 3^rd^, 2019*"
output: ioslides_presentation
---

```{r setup,include=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r data}
#set the working directory before start
setwd("~/Desktop/Data Science @Coursera/Assignments/9_3")
#Download data
if (!file.exists("november_generator2018.xlsx")){
      download.file("https://www.eia.gov/electricity/data/eia860m/xls/november_generator2018.xlsx", 
                    "november_generator2018.xlsx", method="curl")
}
#read data
library(readxl)
generator <- read_excel("november_generator2018.xlsx", sheet = "Operating", skip = 1)
#process to get MA power plants
library(dplyr)
gen <- generator %>% filter(`Plant State`=="MA") %>%
      select(c(`Plant ID`,`Plant Name`,`Plant State`,
               `Nameplate Capacity (MW)`,`Energy Source Code`,
               `Operating Year`, County,Latitude,Longitude)) %>% 
      filter(`Longitude`< -70) %>%
      group_by(`Plant ID`) %>% 
      mutate(`Capacity (MW)`=sum(`Nameplate Capacity (MW)`)) %>%
      select(-`Nameplate Capacity (MW)`) %>%
      distinct(`Plant ID`, .keep_all = TRUE) %>% 
      mutate(Label = ifelse(`Energy Source Code`=="DFO"|
                              `Energy Source Code`=="KER"|
                              `Energy Source Code`=="RFO","Oil",
                        ifelse(`Energy Source Code`=="NG","Gas",
                               ifelse(`Energy Source Code`=="SUN","Solor",
                                      ifelse(`Energy Source Code`=="WND","Wind",
                                             ifelse(`Energy Source Code`=="WAT","Hydro",
                                                    ifelse(`Energy Source Code`=="NUC","Nuclear",
                                                           ifelse(`Energy Source Code`=="MWH","Storage",
                                                                  "Other"))))))))

rm(generator)

```

## MA Power Plant Map
```{r map}
library(leaflet)
#color for levels(gen$Label)
getcolor <- c("blue","black","gray","brown","green","yellow","red","purple")

MAP <- leaflet() %>% addTiles() %>%
      addMarkers(lat = gen$Latitude,lng=gen$Longitude,
                 popup = paste(gen$`Plant Name`,gen$Label,paste(gen$`Capacity (MW)`,"MW",sep=""),sep=" - "),
                 clusterOptions = markerClusterOptions())  %>%
      addCircleMarkers(weight=1,radius = sqrt(gen$`Capacity (MW)`),color = getcolor,
                       lat = gen$Latitude,lng = gen$Longitude) %>%
      addLegend(labels = unique(gen$Label),colors = getcolor)
MAP
```

## Power Plants by Energy Source
Solar panels are popular now, but solar power is distributed with small capacity.
```{r number}
library(plotly)
plot_ly(gen,x=~Label,type="histogram",color = ~factor(Label)) %>% 
      layout(xaxis=list(title="Energy Source"),yaxis=list(title="Number of Plants"))

```

## Power Plants by Energy Source
Even though solar plants win by numbers, but Natural Gas is the main power source.
```{r capacity}
cap <- aggregate(`Capacity (MW)`~ Label,data=gen,sum)
plot_ly(cap,x=~Label,y=~`Capacity (MW)`,type="bar",color = ~Label) %>% 
      layout(xaxis=list(title="Energy Source"),yaxis=list(title="Total Capacity (MW)"))
```

## Power Plants by Operating Year
From this plot we can see the trend of power plants in MA. 
```{r year}
plot_ly(gen,x=~`Operating Year`,y=~`Capacity (MW)`,type="scatter",color=~factor(Label),mode="markers")
```

## Thank you!

### Reference: 
EIA,Preliminary Monthly Electric Generator Inventory (Form EIA-860M),November,2018


