# Reproducible Research Assignment


### Set working environment
```{r setup}
rm(list=ls())
library(ggplot2)
setwd("~/Desktop/Data Science @Coursera/Assignments/5_2")
```
### Load & Preprocess the data
Download the data from website.
```{r download}
if (!file.exists("activity.zip")){
      fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(fileURL, "activity.zip", method="curl")
}
unzip("activity.zip")
file.remove("activity.zip")
```
Load and process the data. 
```{r data, warning=FALSE}
dat <- read.csv("activity.csv",header = TRUE)
dat$steps <- as.numeric(dat$steps)
dat$date <- as.Date(dat$date,"%Y-%m-%d")
```
Reveiw the data when completed.
```{r review}
str(dat)
head(dat)
```
### What is mean total number of steps taken per day?
1.Calculate the total number of steps taken per day.
```{r aggregate}
dailystep <- aggregate(steps~date,dat,sum)
head(dailystep)
```
2.Make a histogram of the total number of steps taken each day.
```{r plot1}
qplot(steps,data = dailystep,bins=10) +
      geom_rug(size=0.2) +
      labs(x="Daily Total Steps",y="Frequency") +
      ggtitle("Histogram of Daily Steps") +
      theme(plot.title = element_text(hjust=0.5))
```

3.Calculate and report the mean and median of the total number of steps taken per day.
```{r summary}
summary(dailystep$steps)
```
###What is the average daily activity pattern?
1.Make a time series plot of the 5-minute interval and the average number of steps. 
```{r plot2}
interstep <- aggregate(steps~interval,dat,mean)
ggplot(interstep,aes(interval,steps)) + 
      geom_line() +
      labs(x="Average Steps of Each Interval",y="Steps") +
      ggtitle("Average Daily Activity Pattern") +
      theme(plot.title = element_text(hjust=0.5))
```

2.Which 5-minute interval contains the maximum number of steps?
```{r max}
interstep[which.max(interstep$steps),]
```
### Imputing missing values
1.Calculate and report the total number of missing values in the dataset.
In the table below, "True"" means the value is missing.
```{r na}
table(is.na(dat$steps))
```
2.Filling in all of the missing values in the dataset.
Here, missing values are filled by mean of the 5-min interval steps into a new dataset.
```{r fill}
dat2 <- dat
for(i in 1:nrow(dat2)){
      if(is.na(dat2$steps[i])==TRUE){
            dat2$steps[i] = interstep$steps[which(interstep$interval==dat2$interval[i])]
      }
}

table(is.na(dat2$steps))
```
3.Make a histogram.Calculate and report the mean and median total number of steps taken per day. 
```{r plot3}
dailystep2 <- aggregate(steps~date,dat2,sum)
qplot(steps,data = dailystep2,bins=10) +
      geom_rug(size=0.2) +
      labs(x="Daily Total Steps",y="Frequency") +
      ggtitle("Histogram of Daily Steps") +
      theme(plot.title = element_text(hjust=0.5))
summary(dailystep$steps)
```
We can see from the result, filling missing values make it a little bit different from the original statistic. After imputing missing data, the mean keeps the same, but the median change from 10765 to 10766.
### Are there differences in activity patterns between weekdays and weekends?
1.Create a new factor variable indicating weekday or weekend day.
```{r weekday}
dat2$day <- weekdays(dat2$date)
for(i in 1:nrow(dat2)){
      if(dat2$day[i] == "Saturday"|dat2$day[i] == "Sunday"){
            dat2$daytype[i] = "weekend"
      }
      else {
            dat2$daytype[i]="weekday"
      }
} 
```
2.Make a panel plot containing a time series plot of average nummber of steps for each 5-min interval for both weekday and weekend day.
```{r plot4}
interstep2 <- aggregate(steps~interval+daytype,dat2,mean)
ggplot(interstep2,aes(interval,steps,color=daytype)) +
      geom_line() +
      facet_grid(daytype~.) +
      labs(x="Average Steps of Each Interval",y="Steps") +
      ggtitle("Activity Pattern of Weekday & Weekend") +
      theme(plot.title = element_text(hjust=0.5))
```
